---
title: "R Practice: Functions"
author: "Braden Charles DeMattei"
format: html
---

# 16.1

# R Practice: Functions

## 16.1.1

`double_it()`
```{r}
double_it <- function(x){
  
  print(x*2)
  
}
```

Testing function
```{r}
double_it(2)
```
Success! The values are as expected.

`exclaim_age()`
```{r}
exclaim_age <- function(age){
  
  print(paste("I am", age, "years old!", sep = " "))
  
}
```

```{r}
exclaim_age(27)
```
Success! It works perfectly.

## 16.1.2

`find_max()`
```{r}
# "find_max() function" #
find_max <- function(value_1, value_2) {
    
    if (value_1 > value_2) {
        return(value_1)
    }
    else if (value_2 > value_1) {
        return(value_2)
    }
}
```

```{r}

find_max(7, 12)

5*find_max(7, 12) #answer should be 60
```

## 16.1.3

Question 4a:

> Run find_max(4, 2, 5) in the Console.

> What happens? What kind of message appears? Is it sufficient? If not, consider adding a warning or error message using warning() or stop(). Remember, use ?function to access the Help page. Add additional logic to the function, as needed.

An error is returned where the 5 is said to be an unused argument. It is sufficient.

Question 4b:

> Run find_max(4,4) in the Console.

> What happens? What kind of message appears? Is it sufficient? If not, consider adding a warning or error message using warning() or stop(). Remember, use ?function to access the Help page. Add additional logic to the function, as needed.

No message appears but no output is generated either. This is not sufficient.

Updating `find_max()` function to include error message when using same value for both `value_1` and `value_2`
```{r}
find_max <- function(value_1, value_2) {
    
    if(value_1 == value_2){
      stop("Values must be different from each other silly!")
    }
  
    if (value_1 > value_2) {
        return(value_1)
    }
    else if (value_2 > value_1) {
        return(value_2)
    }
}
```

Question 4c:

> Run find_max(4, "cow") in the Console.

> What happens? What kind of message appears? Is it sufficient? If not, consider adding a warning or error message using warning() or stop(). Remember, use ?function to access the Help page. Add additional logic to the function, as needed.

Cow is output as the larger value, which doesn't make any sense within the context of the function. An error message needs to be added.

Adding warning message to `find_max()`
```{r}

find_max <- function(value_1, value_2) {
    
    if(value_1 == value_2){
      stop("Values must be different from each other, silly!")
    }
  
    if(is.character(value_1) | is.character(value_2)){
      stop("Values must be numeric, ya dingus!")
    }
  
    if (value_1 > value_2) {
        return(value_1)
    }
    else if (value_2 > value_1) {
        return(value_2)
    }
}
```

